---
title: "TCGA Cancer Sample Data"
author: "Britney Pheng"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
    html:
        fontsize: 14pxs
        toc: true
        top-depth: 3
output: html_document
editor: visual
---

# Initial Setup

## Load packages

```{r packages, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  # BiocManager
  library(GenomicSuperSignature)
  library(curatedTCGAData)
  library(MultiAssayExperiment)
  library(TCGAutils)
  library(ComplexHeatmap)
  
  # CRAN
  library(tidyverse) # includes dplyr, ggplot2, magrittr, tidyr
  library(magick)
  library(wordcloud)
  library(ztable)
  library(metafolio)
  library(randomForest)
  library(caret)
})
```

## Load data

### Load RAVmodel

```{r RAVmodel, warning=FALSE, message=FALSE}
RAVmodel <- getModel('C2', load=TRUE)
```

### COAD: Colon adenocarcinoma

```{r COAD, warning=FALSE, message=FALSE}
coad <- curatedTCGAData(diseaseCode = 'COAD',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

sampleTables(coad) # Display the sample types (e.g., normal, cancer)

coad2 <- TCGAsplitAssays(coad, c("01", "11"))

coad_rna_cancer <- getWithColData(coad2,
                           '01_COAD_RNASeq2Gene-20160128')

assay(coad_rna_cancer) <- log2(assay(coad_rna_cancer) + 1)

coad_meta_cancer <- colData(coad_rna_cancer)
```

```{r}
save(coad_meta_cancer, file = "TCGA_COAD_cancer_dataset.rda")
```

```{r}
validate_coad_cancer <- validate(coad_rna_cancer, RAVmodel)
validate_coad_cancer <- validate_coad_cancer %>% arrange(desc(score))

heatmapTable(validate_coad_cancer, RAVmodel, num.out=10, column_title="COAD Cancer Samples")
```

### LUAD: Lung adenocarcinoma

```{r}
luad <- curatedTCGAData(diseaseCode = 'LUAD',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

luad2 <- TCGAsplitAssays(luad, c("01", "11"))

luad_rna_cancer <- getWithColData(luad2,
                           '01_LUAD_RNASeq2Gene-20160128')

assay(luad_rna_cancer) <- log2(assay(luad_rna_cancer) + 1)

luad_meta_cancer <- colData(luad_rna_cancer)
```

```{r}
save(luad_meta_cancer, file = "TCGA_LUAD_cancer_dataset.rda")
```

### LUSC: Lung squamous cell carcinoma

```{r}
lusc <- curatedTCGAData(diseaseCode = 'LUSC',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

lusc2 <- TCGAsplitAssays(lusc, c("01", "11"))

lusc_rna_cancer <- getWithColData(lusc2,
                           '01_LUSC_RNASeq2Gene-20160128')

assay(lusc_rna_cancer) <- log2(assay(lusc_rna_cancer) + 1)

lusc_meta_cancer <- colData(lusc_rna_cancer)
```

```{r}
save(lusc_meta_cancer, file = "TCGA_LUSC_cancer_dataset.rda")
```

### BRCA: Breast invasive carcinoma

```{r}
brca <- curatedTCGAData(diseaseCode = 'BRCA',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

brca2 <- TCGAsplitAssays(brca, c("01", "11"))

brca_rna_cancer <- getWithColData(brca2,
                           '01_BRCA_RNASeq2Gene-20160128')

assay(brca_rna_cancer) <- log2(assay(brca_rna_cancer) + 1)

brca_meta_cancer <- colData(brca_rna_cancer)
```

```{r}
save(brca_meta_cancer, file = "TCGA_BRCA_cancer_dataset.rda")
```

### KIRC: Kidney renal clear cell carcinoma

```{r}
kirc <- curatedTCGAData(diseaseCode = 'KIRC',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

kirc2 <- TCGAsplitAssays(kirc, c("01", "11"))

kirc_rna_cancer <- getWithColData(kirc2,
                           '01_KIRC_RNASeq2Gene-20160128')

assay(kirc_rna_cancer) <- log2(assay(kirc_rna_cancer) + 1)

kirc_meta_cancer <- colData(kirc_rna_cancer)
```

```{r}
save(kirc_meta_cancer, file = "TCGA_KIRC_cancer_dataset.rda")
```

### KIRP: Kidney renal papillary cell carcinoma

```{r}
kirp <- curatedTCGAData(diseaseCode = 'KIRP',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

kirp2 <- TCGAsplitAssays(kirp, c("01", "11"))

kirp_rna_cancer <- getWithColData(kirp2,
                           '01_KIRP_RNASeq2Gene-20160128')

assay(kirp_rna_cancer) <- log2(assay(kirp_rna_cancer) + 1)

kirp_meta_cancer <- colData(kirp_rna_cancer)
```

```{r}
save(kirp_meta_cancer, file = "TCGA_KIRP_cancer_dataset.rda")
```

### HNSC: Head and neck squamous cell carcinoma

```{r}
hnsc <- curatedTCGAData(diseaseCode = 'HNSC',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

hnsc2 <- TCGAsplitAssays(hnsc, c("01", "11"))

hnsc_rna_cancer <- getWithColData(hnsc2,
                           '01_HNSC_RNASeq2Gene-20160128')

assay(hnsc_rna_cancer) <- log2(assay(hnsc_rna_cancer) + 1)

hnsc_meta_cancer <- colData(hnsc_rna_cancer)
```

```{r}
save(hnsc_meta_cancer, file = "TCGA_HNSC_cancer_dataset.rda")
```

### UCEC: Uterine corpus endometrial carcinoma

```{r}
ucec <- curatedTCGAData(diseaseCode = 'UCEC',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

ucec2 <- TCGAsplitAssays(ucec, c("01", "11"))

ucec_rna_cancer <- getWithColData(ucec2,
                           '01_UCEC_RNASeq2Gene-20160128')

assay(ucec_rna_cancer) <- log2(assay(ucec_rna_cancer) + 1)

ucec_meta_cancer <- colData(ucec_rna_cancer)
```

```{r}
save(ucec_meta_cancer, file = "TCGA_UCEC_cancer_dataset.rda")
```

### OV: Ovarian serous cystadenocarcinoma

```{r}
ov <- curatedTCGAData(diseaseCode = 'OV',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

ov2 <- TCGAsplitAssays(ov, c("01", "11"))

ov_rna_cancer <- getWithColData(ov2,
                           '01_OV_RNASeq2Gene-20160128')

assay(ov_rna_cancer) <- log2(assay(ov_rna_cancer) + 1)

ov_meta_cancer <- colData(ov_rna_cancer)
```

```{r}
save(ov_meta_cancer, file = "TCGA_OV_cancer_dataset.rda")
```

### PAAD: Pancreatic adenocarcinoma

```{r}
paad <- curatedTCGAData(diseaseCode = 'PAAD',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

paad2 <- TCGAsplitAssays(paad, c("01", "11"))

paad_rna_cancer <- getWithColData(paad2,
                           '01_PAAD_RNASeq2Gene-20160128')

assay(paad_rna_cancer) <- log2(assay(paad_rna_cancer) + 1)

paad_meta_cancer <- colData(paad_rna_cancer)
```

```{r}
save(paad_meta_cancer, file = "TCGA_PAAD_cancer_dataset.rda")
```

### PRAD: Prostate adenocarcinoma

```{r}
prad <- curatedTCGAData(diseaseCode = 'PRAD',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

prad2 <- TCGAsplitAssays(prad, c("01", "11"))

prad_rna_cancer <- getWithColData(prad2,
                           '01_PRAD_RNASeq2Gene-20160128')

assay(prad_rna_cancer) <- log2(assay(prad_rna_cancer) + 1)

prad_meta_cancer <- colData(prad_rna_cancer)
```

```{r}
save(prad_meta_cancer, file = "TCGA_PRAD_cancer_dataset.rda")
```

## Create vector dataset

```{r}
TCGA_cancer_validationDatasets <- vector(mode = "list", length = 11)
names(TCGA_cancer_validationDatasets) <- c("BRCA", "COAD", "HNSC", "KIRC", "KIRP", "LUAD",
                                    "LUSC", "OV", "PAAD", "PRAD", "UCEC")

TCGA_cancer_validationDatasets[[1]] <- brca_rna_cancer
TCGA_cancer_validationDatasets[[2]] <- coad_rna_cancer
TCGA_cancer_validationDatasets[[3]] <- hnsc_rna_cancer
TCGA_cancer_validationDatasets[[4]] <- kirc_rna_cancer
TCGA_cancer_validationDatasets[[5]] <- kirp_rna_cancer
TCGA_cancer_validationDatasets[[6]] <- luad_rna_cancer
TCGA_cancer_validationDatasets[[7]] <- lusc_rna_cancer
TCGA_cancer_validationDatasets[[8]] <- ov_rna_cancer
TCGA_cancer_validationDatasets[[9]] <- paad_rna_cancer
TCGA_cancer_validationDatasets[[10]] <- prad_rna_cancer
TCGA_cancer_validationDatasets[[11]] <- ucec_rna_cancer

save(TCGA_cancer_validationDatasets, file = "TCGA_cancer_validationDatasets.rda")
```

### Validate

```{r}
val_all <- validate(TCGA_cancer_validationDatasets, RAVmodel)
```

```{r}
heatmapTable(val_all, RAVmodel, num.out=10)
```
